---
title: "Lab 6 - GeoTIFF, kriging"
author: "Amelia Ritger"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

### Attach packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
#library(tmaptools)
library(gstat)
```

## Read in the data
```{r}
gc_dem <- raster(here("data","gc_dem.tif"))
```

## Go exploring
```{r}
#Plot it
plot(gc_dem)

#Check the CRS
gc_dem@crs

#Check the extent (bounding box) 
gc_dem@extent

#Create a wgs84 with lat/long:
wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

#Reproject
gc_reproj <-  projectRaster(gc_dem, crs = wgs84, method = "bilinear")
gc_reproj@extent
```

### Crop raster to a smaller area (bounding box):
```{r}
bounds <- as(extent(-112.4, -112.0, 36.1, 36.3),'SpatialPolygons')
```

### Make the CRS of our bounding box the same as for gc_reproj:
```{r}
crs(bounds) <- crs(gc_reproj)
```

### Crop our original spatial data using "bounds" polygon:
```{r}
gc_crop <- crop(gc_reproj, bounds)
plot(gc_crop)
```

### Resample using `aggregate()` function:
```{r}
gc_agg <- raster::aggregate(gc_crop, fact = 30)
plot(gc_agg)
```

## Now let's get safe and sound in ggplot:

First, convert data to a df
```{r}
gc_df <- as.data.frame(gc_agg, xy=TRUE)

ggplot(data=gc_df, aes(x=x, y=y)) +
  geom_raster(aes(fill=gc_dem)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c(
    "purple", "magenta", "turquoise","blue", "darkblue"))
```

## How can I just select cells that match a given criteria?

*Let's say vegetables only grow when gc_dem is between 1000 and 1500...*
```{r}
gc_hab <- gc_crop

#Set any cells outside (1000 - 1500) to NA
gc_hab[gc_hab>1500 | gc_hab<1000] <- NA #index it with [], meaning you set conditions

plot(gc_hab)
```

Now let's make this interactive with `tmap`
```{r}
tmap_mode("view")

tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")
```

### It's kriging rain in Kansas!
*now that I know what kriging is, this joke makes no sense*

Read in the KS counties shapefile data
```{r}
ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp"))

#Use base plot:
plot(ks_counties)

#Check the CRS:
st_crs(ks_counties) #NA

#Set the CRS
st_crs(ks_counties) <- 4326
```

Practice in ggplot:
```{r}
ggplot(data=ks_counties) +
  geom_sf()
```

Read in the rainfall data:
```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names()
```

Update the ks_rain data to be recognized as spatial points:
```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs=4326)
```

```{r}
ggplot() +
  geom_sf(data=ks_counties) + #add counties layer
  geom_sf(data=ks_sf, aes(color=amt, size=amt), show.legend = FALSE) #add ks_rain layer
```

## Kriging to predict rainfall:
```{r}
ks_sp <- as_Spatial(ks_sf)
```

### First, make a spatial pixels grid that we'll make predictions over
```{r}
lat <- seq(37, 40, length.out = 200) #creates a sequence of values ranging from 37.0 to 40.0 and there are 200 evenly spaced values between them
long <- seq(-94.6, -102, length.out = 200) #based on bounding box values obtained from bbox(ks_sp)

#Now make this into a spatial grid
grid <- expand.grid(lon=long, lat = lat)

#Tell R this is spatial data
grid_sf <- st_as_sf(grid, coords=c("lon","lat"), crs=4326)

grid_sp <- as_Spatial(grid_sf)
plot(grid_sp)
```

### Now, let's make a variogram
```{r}
ks_vgm <- variogram(amt~1, data=ks_sp)
plot(ks_vgm)
```

Nugget looks like 0.1
Sill looks like... idk 1.0
Range looks like ... idk 200

```{r}
ks_vgm_fit <- fit.variogram(ks_vgm, model=vgm(nugget=0.1, psill=1, range=200, model = "Sph")) #there are tons of different model options, spherical looks like a good fit here

#model=vgm() inputs are pretty resilient as long as you're in the ballpark and not multiple orders of magnitude away from the actual values of nugget, sill, range

plot(ks_vgm, ks_vgm_fit)
ks_vgm_fit
```

### Now, krige!
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)
#ks_krige@data #shows predicted value for rainfall and variance (uncertainty) value

spplot(ks_krige, "var1.pred")
```

### Make a data frame of kriged predictions
```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

#Convert this df to sf object
rain_sf <- st_as_sf(ks_df, coords=c("longitude", "latitude"), crs=4326)

ggplot(rain_sf) +
  geom_sf(aes(color=var1.pred))
```

### Crop this rectangle to the actual KS state outline:
```{r}
thisisks <- read_sf(here("data", "states"), 
              layer="cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME=="Kansas") %>% 
  st_transform(crs=4326)

plot(thisisks)
#Yep, looks like Kansas

#Find the intersection of the KS outline and the kriged grid
rain_sf_ks <- st_intersection(rain_sf, thisisks)

#Plot!
ggplot(data=rain_sf_ks) +
  geom_sf(aes(color=var1.pred)) #geom_raster would make the corners non-rounded edges
```



